---
title: "4. Light pollution - Very complicated"
output:
  html_document:
    toc: yes
    theme: lumen
    highlight: tango
    # highlight: highlight\atom-one-dark.theme
    df_print: paged
---

# Setting up

<!-- install packages -->

```{r install-packages, echo=FALSE}

# These are the packages you already have

packages_you_have <- installed.packages()[, "Package"]

# These are the packages we'll use in this document

packages_we_use <- c("rmarkdown", "jsonlite", "dplyr", "tibble", "tidyr", "readr", "stringr", "leaflet", "leaflet.extras", "leafem", "terra", "stars", "geojsonio", "sf", "lwgeom", "viridisLite", "here", "htmlwidgets", "htmltools", "conflicted")

# These are the packages you need to install

packages_you_need <- packages_we_use[!packages_we_use %in% packages_you_have]

# This will install the packages you don't already have

if (length(packages_you_need)) {
    
    cat("Installing:", paste(packages_you_need, collapse = ", "))
    
    install.packages(packages_you_need)
}

```

## Load packages

```{r load-packages, echo=TRUE, message=FALSE, warning=FALSE}

library(jsonlite)
library(dplyr)
library(tibble)
library(tidyr)
library(readr)
library(stringr)
library(here)
library(fs)
library(leaflet)
library(leaflet.extras)
library(leafem)
library(terra)
library(stars)
library(geojsonio)
library(sf)
library(viridisLite)
library(htmlwidgets)
library(htmltools)
library(conflicted)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

```

## Customizations

### Modified EasyButtons

`addResetMapButton` from `leaflet.extras` but allowing specification of position  
`addEasyButton` from `leaflet`, but removing fontawesome dependency (so I can use the current version from `plugins`)

```{r EasyButtons}

source(here("code/functions/addResetMapButtonPosition.R"))
source(here("code/functions/addEasyButtonNoFaDeps.R"))

```

### Plugins

Define a function to add plugins as map dependencies. These will get inserted into the HTML output with `<script>` (JavaScript) and `<link>` (CSS) tags.

```{r registerPlugin}

registerPlugin <- 
    function(map, plugin) {
        map$dependencies <- c(map$dependencies, list(plugin))
        map
    }

```

Fontawesome is a large set of icons, including lots of [free icons](https://fontawesome.com/icons?d=gallery&m=free). We'll use these to style the markers.

```{r fontawesome}

# my own FA library

fa_dir <- here("code/plugins/@fortawesome/fontawesome-free")

fa_plugin <-
    htmlDependency(
        name = "fontawesome", 
        version = fromJSON(path(fa_dir, "package.json"))$version,
        src = c(file = fa_dir),
        stylesheet = "css/all.css",
        all_files = TRUE
    )

```

`geoblaze` is a JavaScript library for fast raster computations. Here we'll be using it to identify the raster value on click, which we will send to the dark points function.

```{r geoblaze}

# geoblaze raster computation

geoblaze_dir <- here("code/plugins/geoblaze")

geoblaze_plugin <-
    htmlDependency(
        name = "geoblaze", 
        version = fromJSON(path(geoblaze_dir, "package.json"))$version,
        src = c(file = path(geoblaze_dir, "dist")),
        script = "geoblaze.web.min.js",
        all_files = FALSE
    )

```

`ExtraMarkers` is a leaflet extension that allows more control over marker styling.

```{r ExtraMarkers}

ExtraMarkers_dir <- here("code/plugins/Leaflet.ExtraMarkers")

ExtraMarkers_plugin <-
    htmlDependency(
        name = "ExtraMarkers", 
        version = fromJSON(path(ExtraMarkers_dir, "package.json"))$version,
        src = c(file = path(ExtraMarkers_dir, "dist")),
        stylesheet = "css/leaflet.extra-markers.min.css",
        script = "js/leaflet.extra-markers.min.js",
        all_files = TRUE
    )


```

`Control.Geocoder` does what it sounds like: takes an address and geocodes it. The geocoded address will be treated like a map click.

```{r geocoder}

Geocoder_dir <- here("code/plugins/Control.Geocoder")

Geocoder_plugin <-
    htmlDependency(
        name = "geocoder",
        version = fromJSON(path(Geocoder_dir, "package.json"))$version,
        src = list(file = path(Geocoder_dir, "dist")),
        stylesheet = "Control.Geocoder.css",
        script = "Control.Geocoder.js",
        all_files = TRUE
    )

```


### Custom JavaScript and HTML for `onRender`

```{r custom-JS}

closest_dark_place_js <- read_file(here("code/js/closest_dark_place.js"))
dark_point_control    <- read_file(here("code/html/dark_point_control.html"))


# Replacing "####" in JavaScript with HTML

closest_dark_place <- 
    str_replace(
        closest_dark_place_js, 
        "####",
        dark_point_control
    )


```


# Load and process data

## Reading raster

```{r read-raster}

sky_brightness <- 
    rast(here("data/ny_sky_brightness_geotiff.tif")) %>% 
    st_as_stars() %>% 
    rename(brightness_values = ny_sky_brightness_geotiff.tif)

```

### Getting bounding box for setting map bounds

```{r bbox}

sky_brightness_bbox <- st_bbox(sky_brightness)

```

### Converting to a `tbl`, to pass to `onRender`

```{r coords}

sky_brightness_coords <- 
    sky_brightness %>% 
    as_tibble() %>% 
    drop_na(brightness_values)

```

# Constructing the map

We'll first initialize the map, then add tiles[^1], add the light pollution raster, add some map control, add an image query that tells us the raster value under the mouse pointer, add plugins and dependencies, and finish by adding our custom JavaScript using `onRender`.

```{r map}

# access token for JawgMaps (free, and can be revoked!)

jawg_token <- "jBn4H6Bv04xoEkuaRdWm4vIcIJjGYmLsD1jZ2kRL5uSZk61d1YhwzvdVM4FBaadM"


light_pollution_map <- 
    
    # initialize
    
    leaflet(
        options = list(
            "duration" = 0.375,
            "zoomSnap" = 0.5,
            "padding" = c(10, 10),
            "preferCanvas" = FALSE, 
            "updateWhenZooming" = FALSE,
            "updateWhenIdle" = TRUE
        )
    ) %>%
    
    fitBounds(
        lng1 = sky_brightness_bbox[[1]],
        lat1 = sky_brightness_bbox[[2]],
        lng2 = sky_brightness_bbox[[3]],
        lat2 = sky_brightness_bbox[[4]]
    ) %>%
    
    # enable browser tile caching, to speed up the map
    
    enableTileCaching() %>%
    
    # add tiles in groups, which we'll use for the tile layer switcher
    
    # zIndex = -1000 makes sure that the tiles are always the lowest layer
    
    addProviderTiles(
        # provider = providers$Stadia.AlidadeSmoothDark, 
        # provider = providers$CartoDB.DarkMatter, 
        provider = providers$Jawg.Matrix, 
        options = providerTileOptions(zIndex = -1000, accessToken = jawg_token),
        group = "Dark"
    ) %>%
    addProviderTiles(
        # provider = providers$Stadia.StamenTonerLite, 
        # provider = providers$CartoDB.Positron, 
        provider = providers$Jawg.Light, 
        options = providerTileOptions(zIndex = -1000, accessToken = jawg_token),
        group = "Light"
    ) %>%
    addProviderTiles(
        provider = providers$CartoDB.Voyager, 
        options = providerTileOptions(zIndex = -1000),
        group = "Streets"
    ) %>%
    
    # you xcan also add tiles from other places
    
    addTiles(
        urlTemplate = 
            paste0(
                "//services.arcgisonline.com/ArcGIS/rest/services/",
                "USA_Topo_Maps/MapServer/tile/{z}/{y}/{x}"
            ), 
        attribution = 
            paste0(
                "Map tiles by <a href='http://goto.arcgisonline.com/maps/USA_Topo_Maps'>Esri</a> - ",
                "Map Data © 2013 National Geographic Society, i-cubed"
            ), 
        options = tileOptions(zIndex = -1000),
        group = "Topo"
    ) %>% 
    
    # sky brightness raster
    
    addGeoRaster(
        x = sky_brightness,
        project = TRUE,
        group = "Sky Brightness",
        layerId = "Sky Brightness",
        resolution = 64,
        colorOptions =
            colorOptions(
                palette = inferno(64, direction = -1),

                # modifying breaks to get a better mapping of visual differences to
                #   photometric categories
                #
                # 16 = 2^4, so need 4 `sqrt()` calls to reverse:

                breaks =
                    sqrt(sqrt(sqrt(sqrt(
                        seq(
                            min(sky_brightness$brightness_values, na.rm = TRUE)^16,
                            max(sky_brightness$brightness_values, na.rm = TRUE)^16,
                            length.out = 64
                        )
                    )))),
                na.color = "#00000000"
            ),
        options =
            tileOptions(
                zIndex = 1000,
                updateWhenZooming = FALSE,
                updateWhenIdle = TRUE
            )
    ) %>%
    
    # adding controls
    
    # all the `group` and `layerId` arguments have to be the same to get the raster, image query, and layers control to work (as far as I can tell)

    addLayersControl(
        baseGroups = c("Dark", "Light", "Streets", "Topo"),
        overlayGroups = "Sky Brightness",
        options = layersControlOptions(collapsed = FALSE, autoZIndex = FALSE),
        position = "topright"
    ) %>%
    
    # raster mouseover values (put this after layers control for better positioning on map)
    
    addImageQuery(
        x = sky_brightness,
        group = "Sky Brightness",
        layerId = "Sky Brightness",
        position = "topright",
        digits = 2,
        type = "mousemove",
        prefix = "",
        project = TRUE
    ) %>% 
    
    # reset buttons
    
    addResetMapButtonPosition(position = "bottomleft") %>%
    
    # registering dependencies
    
    addAwesomeMarkersDependencies(libs = c("ion", "glyphicon")) %>%

    # registering plugins
    
    registerPlugin(fa_plugin) %>%
    registerPlugin(geoblaze_plugin) %>%
    registerPlugin(ExtraMarkers_plugin) %>%
    registerPlugin(Geocoder_plugin) %>%
    
    # adding specialty JavaScript to find closest dark place to click
    
    onRender(
        str_c(
            "function(el, x, data) {\n",
            closest_dark_place,
            "}"
        ),
        data = sky_brightness_coords
    )

light_pollution_map

```

# Save the map

We'll save 2 different versions: self-contained, and non-self-contained

## Self-contained

In this version, all the data and dependencies are included in-line in the single HTML output file.

```{r save-self-contained}

saveWidget(
    widget = light_pollution_map,
    file = here("output/4-light-pollution_self-contained.html"),
    selfcontained = TRUE,
    title = "4. Light pollution - Very complicated"
)

```

The self-contained version can get pretty big, but you don't have to worry about anything except that 1 output file.

## Non-self-contained

Here, the data and dependencies are put into folders, which the HTML will point to.

```{r save-non-self-contained}

saveWidget(
    widget = light_pollution_map,
    file = here("output/4-light-pollution_non-self-contained.html"),
    selfcontained = FALSE,
    title = "4. Light pollution - Very complicated"
)

```


[^1]: You can check out the provider tiles with the [Leaflet-providers preview](https://leaflet-extras.github.io/leaflet-providers/preview/)
